DATA 621 - Business Analytics and Data Mining

### 2020-12-04 Final_project/Final Project Paper - Group 2.Rmd
Plot the two created features and update the text to match the new results

# Methodology
Using the famous data set from Kaggle [@Mitchell2019TYV], we will explore
relationships between a video’s views and the number of likes, dislikes, and
comments using R [@RCT2020RAL]. We may also use a video’s country of origin and
the category or type of video as predictors.

@@

We will add two features. The first is `disp`: the disproportion between the
likes and dislikes. To minimize division by 0 errors, it is defined as follows:
\[
disp =\begin{cases}
\mathrm{ratings\;enabled}\quad\frac{\min(likes, dislikes)}{\max(likes, dislikes)}\\
\mathrm{ratings\;disabled}\quad 1
\end{cases}
\]
This constrains the ratio to \([0 \leq disp \leq 1]\) with a maximum when the
two are equal, and division by 0 only when both are 0. This usually occurs when
ratings are disabled. There is only `r ldOn` case out of the `r nobsUSExt`
observations where the ratings were not disabled, yet there are neither likes
nor dislikes. As this is a distinct incongruity for YouTube, we will remove that
one observation from the data. When ratings are disabled, perforce there is no
disparity so the ratio will be set to 1.

@@

# Add the engineered feature
usExtract[, `:=`(disp = ifelse(rtgDisabled, 1,
                               pmin(likes, dislikes) / pmax(likes, dislikes)),
                 exceedMean = factor(ifelse(views > usMean, 1, 0),
                                     levels = c(0, 1), labels = c(0, 1)))]
# Melt for exceedMean
usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE,
               variable.name = 'measure', value.name = 'value')

@@

This model will start with the numeric and logical features and the engineered
disparity ratio. It will not use the exceeding mean indicator, as that is
generated from the target variable, and it is felt it will distort the
prediction. It will then proceed through feature selection using the AIC as the
optimization metric.

@@

The signs of these coefficients make sense in the main. Increased `likes` and
`dislikes` are correlated with increased views. Of course one usually views a
video at least once prior to rating it. Increased `comments` is negatively
correlated with views.

We set the baseline category to `Entertainment`, considering it was the most
popular. Therefore we expected negative coefficients for other categories found
significant. We were surprised that `Film & Animation` was an exception. The
"worst" category predictor by magnitude is the `Nonprofits & Activism`.

@@

What also seems counterintuitive at first are the negative correlations between
`lkratio` and `unan` being true with views. It may well be that conflict is good
for Youtube videos. Perhaps the more a video is talked about or argued over, the
more views it will get!

@@

This model has more predictors than does the simple Gaussian (G)LM, although
their signs are similar. What stands out is how many more categories were
considered significant.

### ElasticNet: Penalized Regression
Instead of using AIC to select features, one can make use of penalized
regression. Using an \(L_1\) penalty is the underpinnings of Lasso regression,
which can perform feature selection. Using a squared error term, \(L_2\), is at
the heart of ridge regression [@HTF2009TEo]. Using both methods together is
called the elastic net [@ZH2005RaV; ZH2020eEN]. To tune the hyperparameters,
which includes the weighting between Lasso and Ridge, we will use 10-fold
cross-validation.

@@

Nevertheless, the *relative* magnitude and sign of the parameters are in line
with our expectations. `Nonprofits & Activism` has the largest factor effect, to
the worse. Both `likes` and `dislikes` are positively correlated with views and
`comments` is negatively correlated. A few more categories join
`Film & Animation` as contributing to excess views over the baseline
`Entertainment`. Lastly, having a preponderance of likes tends to reduce views.
Conflict seems to be good for Youtubers!

@@

```{r loadLibraries}
```
```{r loadData}
```
<!-- Data Exploration -->
```{r dataPrep}
```
```{r numLogDensities}
```
```{r catPlot}
```
```{r subsetExtraction}
```
```{r likeDislikeCheck}
```
```{r addFeatures}
```

>>>>

# Methodology
Using the famous data set from Kaggle [@Mitchell2019TYV], we will explore
relationships between a video’s views and the number of likes, dislikes, and
comments using `R` [@RCT2020RAL] and the `caret` package [@Kuhn2020cCa]. We may
also use a video’s category as predictors.

@@

We will add two features. The first is `balance`: the ratio between the
likes and dislikes. The hypothesis is that if a video is either universally
loved or panned, it will get fewer views than if there is a healthy disagreement
about it.

To minimize division by 0 errors, `balance` is defined as follows:
\[
balance =\begin{cases}
\mathrm{ratings\;enabled}\quad\frac{\min(likes, dislikes)}{\max(likes, dislikes)}\\
\mathrm{ratings\;disabled}\quad 1
\end{cases}
\]
This constrains the ratio to the interval \([0, 1]\) with a maximum of 1 when
the two are equal. Now, division by 0 can only occure when both are 0. This 
usually occurs when ratings are disabled. There is only `r ldOn` case out of the
`r nobsUSExt` observations where the ratings were not disabled, yet there are
neither likes nor dislikes. As this is a distinct incongruity for YouTube, we
will remove that one observation from the data. When ratings are disabled,
perforce there is no disparity so the ratio will be set to 1.

@@

# Add the engineered feature
usExtract[, `:=`(balance = ifelse(rtgDisabled, 1,
                               pmin(likes, dislikes) / pmax(likes, dislikes)),
                 exceedMean = factor(ifelse(views > usMean, 1, 0),
                                     levels = c(0, 1), labels = c(0, 1)))]
```
```{r plotBalance, fig.height=4, fig.width=8}
ggplot(usExtract[, .(balance)], aes(x = balance)) +
  geom_density() +
  ggtitle("Distribution of Balance")
```

It's pretty clear from the distribution of `balance` that, at least for videos
viewed in the US, there is a healthy dose of disagreement!

```{r plotExceedMean, fig.height=4, fig.width=8}
# Melt for exceedMean
usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE,
               variable.name = 'measure', value.name = 'value')

@@

This model will start with the numeric and logical features and the engineered
disparity ratio. It will not use the exceeding mean indicator, as that is
generated from the target variable, and it is felt it will distort the
prediction to regress `views` on a function of `views`. The algorithm will
proceed through feature selection using the AIC as the optimization metric.

@@

The signs of these coefficients make sense in the main. Increased `likes` and
`dislikes` are correlated with increased views. Of course one usually views a
video at least once prior to rating it. Interestingly, increased `comments` are
negatively correlated with views.

We set the baseline category to `Entertainment`, considering it was the most
popular. Therefore we expected negative coefficients for other categories found
significant. We were surprised that `Film & Animation` was an exception. The
"worst" category predictor by magnitude is clearly `Nonprofits & Activism`.

@@

When ratings are disabled, there is a bump to views. This is probably because
the intercept is artificially low due to the predictive power of `likes` and
`dislikes`. To balance that when there are none needs a big boost.

What may be most interesting is that the disparity ratio indicator us a very
powerful and significant indicator. Conflict seems to be good for Youtube
videos. The more a video is argued over, the more views it seems to get!

@@

This model has more predictors than does the simple Gaussian (G)LM. Moreover,
the signs are different from the simple LM. The very fact that
`Nonprofits & Activism` has a positive coefficient should raise concerns. This
model is probably a poor one.

### ElasticNet: Penalized Regression
Instead of using AIC to select features, one can make use of penalized
regression. Using an \(L_1\) penalty is the underpinnings of Lasso regression,
which can perform feature selection. Using a squared error term, \(L_2\), is at
the heart of ridge regression [@HTF2009TEo]. Using both methods together is
called the elastic net [@ZH2005RaV; @ZH2020eEN]. To tune the hyperparameters,
which includes the weighting between Lasso and Ridge, we will use 10-fold
cross-validation.

@@

Nevertheless, the *relative* magnitude and sign of the parameters are in line
with our expectations. Both `likes` and `dislikes` are positively correlated
with views and `comments` is negatively correlated. A few more categories join
`Film & Animation` as contributing to excess views over the baseline
`Entertainment`. But `Nonprofits & Activism` has the largest factor by far, and
it is negative, which stands in contrast to the second model. Lastly, having a
larger disparity (ratio of likes to dislikes closer to 1of likes tends to reduce views.
Conflict seems to be good for Youtubers!

@@

```{r loadLibraries}
```
```{r loadData}
```
<!-- Data Exploration -->
```{r dataPrep}
```
```{r numLogDensities}
```
```{r catPlot}
```
```{r subsetExtraction}
```
```{r likeDislikeCheck}
```
```{r addFeatures}
```
```{r plotBalance}
```
```{r plotExceedMean}
```
